Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-17 paper/paper.tex
added result discussion

\documentclass{sig-alternate}
\usepackage{algorithm,algpseudocode}

@@

\section{Related Work}
General policy and rule-based approaches to dynamic provisioning (e.g. Amazon Auto Scaling\footnote{http://aws.amazon.com/autoscaling} and RightScale\footnote{http://www.rightscale.com}).

@@

\subsection{Static Provisioning Dynamic Scheduling (SPDS)}
\label{sec:spds}
This is the simplest strategy based on static provisioning of resources and
on-line scheduling of workflow tasks. Given the budget $B$ and deadline $d$
together with the hourly price $p$ of a VM it is possible to compute the number
of VMs $N_{VM}$ to provision so that they consume the whole budget before the
deadline.

@@

Initially all the
released tasks from all the workflows are submitted to the queue $Q$. Then, they
are inserted into priority queue $P$ based on the priority of workflows. If
there are free VMs available, the tasks from the priority queue are submitted to
the randomly choosen VMs. The process is repeated each time a task finishes.
Using the priority queue guarantees that tasks from lower priority workflows are
always submitted later than the higher priority ones and that the lower priority
tasks can occupy idle VMs. However, as there is no preemption, long-running
low-priority tasks may delay execution of high-priority tasks when such ones
become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
SPSS algorithm. It can be seen that tasks from lower priority workflows can fill
idle time of provisioned virtual machines.

\begin{figure}[htb] 
\centering
\includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
 \caption{Example of schedule generated using SPSS algorithm: tasks are labelled
 by colors depending on the workflow. }
\label{fig:spss-example}
\end{figure}

@@

\subsection{Scenarios}

\subsection{Results}

\begin{figure*}[htb] 
\centering
\includegraphics[width=1.0\textwidth]{figures/MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows}
\label{fig:}
\end{figure*}

\begin{figure*}[htb] 
\centering
\includegraphics[width=1.0\textwidth]{figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/score-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/score-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. Score is computed as $\sum(1/(p+1))$ where
priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
priority).}
\label{fig:}
\end{figure*}

\begin{figure*}[htb] 
\centering
\includegraphics[width=1.0\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/score2-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
priority).}
\label{fig:}
\end{figure*}

\begin{figure*}[htb] 
\centering
\includegraphics[width=1.0\textwidth]{figures/size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/size-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/size-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/size-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. Score is equal to the total work completed,
i.e. it is sum of runtimes (in hours) of tasks of completed workflows.}
\label{fig:}
\end{figure*}

\begin{figure*}[htb] 
\centering
\includegraphics[width=1.0\textwidth]{figures/cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/cost-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=1.0\textwidth]{figures/cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. Plot shows effective computing cost in \$ per
hour calclated by dividing sum of task runtimes by the total budget. Higher
cost results from lower resource utilization.}
\label{fig:}
\end{figure*}

>>>>

\documentclass{sig-alternate}
\usepackage{algorithm,algpseudocode}
\usepackage{url}

@@

\section{Related Work}
General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
Auto Scaling\footnote{\url{http://aws.amazon.com/autoscaling}} and
RightScale\footnote{\url{http://www.rightscale.com}}).

@@

\subsection{Static Provisioning Dynamic Scheduling\\
(SPDS) }
\label{sec:spds}
This is the simplest strategy based on static provisioning of resources and
on-line scheduling of workflow tasks. Given the budget $B$ and deadline $d$
together with the hourly price $p$ of a VM it is possible to compute the number
of VMs $N_{VM}$ to provision so that they consume the whole budget before the
deadline.

@@

Initially all the
released tasks from all the workflows are submitted to the queue $Q$. Then, they
are inserted into priority queue $P$ based on the priority of workflows. If
there are free VMs available, the tasks from the priority queue are submitted to
the randomly choosen VMs. The process is repeated each time a task finishes.
Using the priority queue guarantees that tasks from lower priority workflows are
always submitted later than the higher priority ones and that the lower priority
tasks can occupy idle VMs. However, as there is no preemption, long-running
low-priority tasks may delay execution of high-priority tasks when such ones
become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
SPDS algorithm. It can be seen that tasks from lower priority workflows can fill
idle time of provisioned virtual machines.

\begin{figure}[htb] 
\centering
\includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
 \caption{Example of schedule generated using SPDS algorithm: tasks are labelled
 by colors depending on the workflow. }
\label{fig:spss-example}
\end{figure}

@@

\subsection{Scenarios}
\label{sec:scenarios}

We have prepared test scenarios to evaluate the proposed algorithms under
such a range of parameters that allows to observe the interesting properties of
algorithms for smaller and larger constraints. For a given ensemble we can
expect in general that the number of work done (in terms of workflows completed)
mainly depends on the budget constraint, which determines the number of
machine-hours that can be provisioned. The ideal but trivial case is then a
deadline long enough to allow running all the workflow tasks sequentially: this
eliminates all the overheads of parallel execution and leads to the highest
resource utilization. More interesting are the cases when the deadline is
shorter, since it forces provisioning of multiple VMs and thus increases
overheads due to parallel execution so the budget cannot be spent so
efficiently.

We assumed that the interesting number of workflows in ensemble should be in the
order of between 10 and 100 workflows. This is motivated by several reasons:
first, such numbers are typical for the real applications, e.g. the number of geographical sites for
CyberShake workflow is in the order of 100. Second, the smaller ensembles
consisting of just a few workflows can be aggregated in a single workflow so
there is no need to treat them as ensembles. Similarly, when the number of
workflows is larger that 100 and each workflow has a large number of parallel
tasks, the problem of efficiently allocating them to the resources becomes much
easier and similar to the bag-of-tasks problem.

\begin{figure}[htb] 
\centering
\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
\caption{Pareto-like distribution of workflow sizes in ensemble. }
\label{fig:ensemble-distribution}
\end{figure}

In order to evaluate the algorithms on a standard ser of workflows, we used
ensembles constructed from the DAGs available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
in the workflow generator gallery~\cite{Bharathi08}. The gallery contains
workflows of five applications from Pegasus and from Indiana
University~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
workflows of different sizes based on a given distribution. Two distributions
are considered:
\begin{itemize}
  \item ensemble of workflows of equal size (constant)
  \item Pareto-like distribution.
\end{itemize}

Pareto-like distribution gives a small number of workflows with large size and a
large number of workflows with small size, as seen in
Fig.~\ref{fig:ensemble-distribution}. The number of largest
workflows (of size $\geq$ 900) is slightly increased to represent the
``long-tail'', i.e. that there is a non-insignificant number of large workflows
in the ensemble. The workflows are sorted according to their size and the largest workflows are
given the highest priority.

In the tests we we assumed that all the VMs are homogeneous and have the
processing speed of 1000 MIPS (million instructions per second), the price for
one VM-hour is \$1 and that the sizes of tasks in the workflow gallery are given
in seconds measured on 1000 MIPS machines. In this study we do not take into
account the heterogeneity of the infrastructure since we assume that it is
always possible to select a VM type that has the best price to performance ratio
for a given application.

For each application type, we selected such ranges of constraints (deadline,
budget) that it is possible to observe the possibly broad result space: from the
tight constraints, when only a small number of workflows can finish to the
liberal constraints when all or almost all the workflows can finish. These
parameters vary among the applications. Montage and CyberShake benchmarks
workflows required the budget in \$20 -- \$150 range and deadlines from 1 to
20 hours to complete the ensemble. LIGO and SIPHT workflows have more long
tasks, so they require budgets up to \$1,500 and deadlines extended to 40
hours. Epigenomics workflow is even more costly and needs approximately \$12,000
and 1500 hours to complete the full ensemble.

\subsection{Metrics}

There are multiple ways to evaluate the algorithms and also different metrics
are relevant to the application users that are needed to take the planning
decisions. The set of metrics for each ensemble run we prepared includes:
\begin{itemize}
  \item number of workflows completed $N_c$,
  \item user preference score (utility function) $s$,
  \item total amount of computing time completed $T_C$,
  \item effective cost of computing per hour $C_{eff}$, which includes
  overheads.
\end{itemize}

\begin{figure*}[t] 
\centering
\includegraphics[width=0.19\textwidth]{figures/MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows}
\label{fig:number-complete-pareto}
\end{figure*}

The number of workflows completed is a simplest metric and allows distinguishing
which algorithm performs better. However, as the workflows have
different priorities and sizes, a less efficient algorithm may schedule smaller
low-priority workflows first, thus increasing the total number of workflows
complete. It may also happen that too conservative admission algoritm may reject
overestimate the costs of some large workflows, thus rejest some high-priority
ones which actually could fit into the schedule. Therefore we defined a user
preference score as:

\begin{equation}
\label{eq:score}
s = \sum_{DAGs\ completed}(1/(p+1))
\end{equation}

where priorities of subsequent workflows in the ensemble are $p=0,1,2,\ldots$ (0
means highest priority). This scoring function gives more weight to the top high
priority workflows, whereas the low priority ones are trated almost equally.

Another metric is the total actual computing time completed and it is calculated
as the sum of all the runtimes of all the tasks of completed workflows. As in
our study we assumed that the resource cost pre hour is equal to \$1, the total
computing time can be easily compared to the allocated budget $b$. Their ratio
$u = (T_C * \$1)/b$ is the resource utilization and its inverse is the effective
resource cost: $C_{eff} = b/T_C$ in dollars per hour.

\subsection{Discussion of Results}

\begin{figure}[b!] 
\centering
%\includegraphics[width=0.19\textwidth]{figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/score-LIGO-n-1000-8-dagh1-40m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/score-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of algorithms based on score, which  is computed as
$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
priority).}
\label{fig:score}
\end{figure}

We run the simulations for the constraint values ranging between the
estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
ca. 50-150 (budget, deadline) points for each application. The runs were
repeated for Pareto-distributed and costant workflow sizes. From this result
space we selected the most interesting plots which give the good overview  of
the observed algorithm behavior.

\begin{figure*}[htb]  
\centering
\includegraphics[width=0.19\textwidth]{figures/size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/size-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/size-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/size-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. Score is equal to the total work completed,
i.e. it is sum of runtimes (in hours) of tasks of completed workflows.}
\label{fig:total-time}
\end{figure*}

\begin{figure*}[b!] 
\centering
\includegraphics[width=0.19\textwidth]{figures/cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/cost-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Effective computing cost in \$ per hour calclated by dividing sum of
task runtimes by the total budget. Higher cost results from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows.}
\label{fig:cost}
\end{figure*}

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values per each application ensemble. The budget values
selected correspond to two opposite cases: {\em small budget} (top) and {\em
large budget} (bottom). The first one allows completing only a small number of
workflows from the ensemble, while the latter one suffices to finish almost all
the workflows. It can be seen that the static planning algorithm (SPSS) performs
better for tighter constraints: when the budget is low to admit only a small
number of workflows, the static procedure is able to more efficiently pack these
workflow tasks onto the resources. This can be clearly seen for the LIGO,
Epigenomics and SIPHT ensembles for small budgets and also for the same
ensembles with large budgets but short deadlines. The exception are Montage and
CyberShake ensembles, which consist of many fine-grained tasks that can be
executed in parallel. For these cases, as well for the other ensembles with
large budgets, the simple dynamic scheduling techniqies perform equally well or
better. The reason for that behavior is that the large number of tasks
guarantees that resources are almost never idle, while the provisioner algorithm
makes sure that the VM instances are most efficiently utilized before shutdown.
We can also observe that the workflow-aware dynamic algorithm (WA-DPDS) performs
better that the workflow-unaware version. This means that for the tested
workflows the simple admission procedure based on estimation of workflow cost
does not degrade the solution, but rather in many cases it allows rejecting
larger workflows that would lead to budget overrun. Thus it can save the space
for smaller workflows that can complete.

Fig.~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
computed for two sample applications. Comparing
Fig.~\ref{fig:number-complete-pareto} to Fig.~\ref{fig:score} we can see that
the latter one gives more flattened shape: this means that e.g. the improvement
of the solution by one workflow may mean adding a one low-priority workflow.
Nevertheless we can see that the curves on both figures have a very similar
shape, which mens that our algorithms obey the priority rules and do not try to
admit lower-priority workflows at the cost of the more important ones.

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric $T_C$. These plots confirm our previous observations about the
algorithms performance, but they also allow to observe a ore general property.
For a given budget, when the deadline is tightened the number of work done
steeply decreases. Assiging a shorter deadline means that more VMs need to be
provisioned to complete the work, but this leads to higher parallelism and thus
to the lower parallel efficiency. E.g. running 2 VMs for 10 hours will be more
efficient than running 10 VMs for 2 hours. Therefore, if deadline permits and
costs are important, parallelism should be avoided when possible.

These observations are further confirmed when examining the efffective resource
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increase of actual costs of computing. This increase reaches ca.
20\%, which means that if we need to run e.g. a SIPHT workflow ensemble with a
short deadline, ca. \$200 out of \$1000 is spent on idle resources.

Another general concluson can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We can see that the selection of deadline and cost can be formulated as
multi-criteria optimization problem and the curves presented in
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the number of work done, whcich direcly results in minimization of effective cost parameter for a given
deadline. Therefore, the obtained results can be used to assist the decison in
resource allocation planning, when both cost and time criteria are given not as
constraints but as objectives.

% \begin{figure*}[htb] 
% \centering
% \includegraphics[width=0.19\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-SIPHT-n-1000-8-dagh5-50m0.pdf}
% \caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
% Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
% priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
% priority).}
% \label{fig:}
% \end{figure*}

\begin{figure}[htb] 
\centering
\includegraphics[width=0.6\columnwidth]{figures/budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}
\caption{Comparison of number of workflows completed for a given deadline,
plotted versus different budget amounts.}
\label{fig:}
\end{figure}
